---
title: "Refformatting rCUX for SYNTAX"
output: html_notebook
---

One script to reformat the db from RCRUX so it is  compatible with SYNTAX 

```{r}
library(tidyverse)
library(eDNAfuns)
```


El objetivo es
`>AB008314;tax=d:Bacteria,p:Firmicutes,c:Bacilli,o:Lactobacillales,
  f:Streptococcaceae,g:Streptococcus;`
  
```{r}
RCRUX <- fasta_reader(here("../../ddbb/CO1_combined/derep_and_clean_db/CO1_combined_derep_and_clean.fasta")) 

RCRUX_tax <- vroom::vroom(file= here("../../ddbb/CO1_combined/derep_and_clean_db/CO1_combined_derep_and_clean_taxonomy.txt"), col_names = c("header", "tax"), delim = "\t") 

RCRUX_tax 

taxonomy_cols <- c("k", "p", "c", "o", "f", "g", "s")

RCRUX_tax |> 
  separate(tax, into= taxonomy_cols, sep = ";") |> 
  pivot_longer(-header, names_to = "rank", values_to = "value" ) |> 
  unite(rank, value, sep = ":", col= "taxa") |> 
  mutate(header2="taxa") |> 
  pivot_wider(id_cols = header, values_from = taxa, names_from =header2, values_fn = {~paste(.x, collapse = ",")} ) |> 
  mutate(taxa= paste0(taxa, ";")) |> 
  inner_join(RCRUX) |> 
  unite(header, taxa, col= "header", sep = ";tax=") |> 
  fasta_writer(header = header, sequence = seq, file.out = "data/rCRUX_for_SYNTAX.fasta")


```

  
  

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
