---
title: "Comparing the length of two sets of sequences"
author: "Ramón Gallego"
date: "`r Sys.Date()`"
output: html_document
params: 
  file1:
    value: ~/Projects/Alumnado/eDNA_jesus_collado/data/UNOISE_March_2025/seqlens_2mismatches.txt
    input: file
  name1:
    value: 2mismatches
    input: text
  file2:
    value: ~/Projects/Alumnado/eDNA_jesus_collado/data/UNOISE_March_2025/seqlens_10mismatches.txt
    input: file
  name2:
    value: default
    input: text    
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning  = FALSE)
```

## Comparing lengths distribution

Given two sets of fasta or fastq files, one might be interested in the lengths distributions. You need to calculate those with `awk`, with the following code

```
awk 'NR%2==1 {print FILENAME, length}' <fasta set 1> | sort | uniq -c > <summary_file_1>

awk 'NR%4==2 {print FILENAME, length}' <fastq set 1> | sort | uniq -c > <summary_file_1>

```
Using awk is not mandatory, you can use your own way of generating a summary file with the columns

Times | FILENAME | Length

Although without a colname, you wuill add it here

```{r}
library(tidyverse)

```

### Reading input files

```{r}
set1 <- read_table(params$file1, col_names = c("times", "filename", "length")) |> 
  mutate(topic = params$name1)
set2 <- read_table(params$file2, col_names = c("times", "filename", "length")) |> 
  mutate(topic = params$name2)
```

```{r}
set1 |> 
  bind_rows(set2) |> 
  ggplot(aes(x = length, y = times)) +
  geom_col()+
  facet_wrap(~topic, ncol = 1)
```

