---
title: "Parsing BLAST"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

We have successfully BLASTed our sequences and now we want to recover the taxonomical information associated with each search

```{r}
library(tidyverse)
library(insect)
library(taxonomizr)
```

## DO it once: get the insect::taxonomy

```{r, eval=FALSE}
taxonomy.now <- insect::taxonomy()

write_rds(taxonomy.now, "../data/taxonomy.rds")
```



## Load the blast_output

```{r}

#taxonomy.now <- read_rds("../data/taxonomy.rds")
taxonomy.now <- read_rds("taxonomy.rds")

#blast_output <- "../pipeline_output/hash_key_221103.txt"
blast_output <- "hash_key_221103.txt"

BLAST_results <- read_table(blast_output, col_names = c("qseqid", "sseqid",  "pident", "length" ,"mismatch", "gapopen" ,"qstart","qend", "sstart", "send" ,"evalue" ,"bitscore", "staxid", "qlen"))


# Many taxids will be repeated, do it only once

BLAST_results %>%
  separate_rows(staxid, sep = ";") -> BLAST_results

BLAST_results %>%
distinct(staxid) -> ALL.TAXIDS
# Bring in those for which we already know

previous.lineages <- read_csv("../data/taxonomies_matches.csv")

# Work only in the new ones

ALL.TAXIDS %>%
  rename(taxID = 1) %>%
  mutate(taxID = as.numeric(taxID)) %>%
  anti_join(previous.lineages) %>%
   mutate(lineage = map(taxID, insect::get_lineage, taxonomy.now)) -> all.lineages


# # Find ids with 2 or more, it takes time in the server
# 
# ALL.TAXIDS %>%
#   filter(! str_detect(staxid, ";")) %>%
#   mutate(staxid = as.numeric(staxid)) %>%
#   mutate(lineage = map(staxid, insect::get_lineage, taxonomy.now)) -> all.lineages

# Save this object
# ALL.TAXIDS = 3872

write_rds(all.lineages, glue::glue("../data/lineages_{Sys.Date()}.rds"))

# Make them into a list
all.lineages %>%
  pull(lineage) -> all.lineages.list

set_names(all.lineages.list, nm= all.lineages$staxid) -> all.lineages.list

# Remove entries without taxonomy
all.lineages.list %>% discard(function(x) is.na(x[[1]]) ) ->  all.lineages.list.good

  all.lineages.list.good %>%
  map(~ bind_cols(.x) %>% mutate (rank = names(.x)) ) %>%
  bind_rows(.id = "taxID") %>%
     filter (rank %in% c( "kingdom", "phylum", "class", "order", "family","genus" , "species")) -> all.lineages.long


  ## Rejoin with previous lineage
  previous.lineages %>%
    bind_rows(all.lineages.long %>% mutate (taxID = as.numeric(taxID))) %>%

write_csv( file.path("../data/taxonomies_matches.csv"))
```

```{r}
custom.lca <- function (df, cutoff = 90) {df %>%  # this function allows to change cutoff parameters for a specified dataframe (df)
  group_by(qseqid) %>%
  select( pident, kingdom, phylum, class, order, family, genus, species) %>%
  nest() %>% # for each query, calculate the agreed taxonomy
  # ungroup %>% slice (1:10) %>%
  mutate(consensus = purrr::map(data,  function(.x) {
    # If there are 100% matches - keep those and calculate the LCA
   
    if (max(.x$pident == 100 )){
       .x %>%
        filter(pident == 100) %>%
        select(-pident) %>%
        condenseTaxa() %>% # agreement in Phylogeny
      paste(., collapse = "%")
      
    }else{
       # If there are no 100% matches, then keep things better than our cutoff
    if(max(.x$pident > cutoff )){

      .x %>%
        filter(pident > cutoff) %>%
        select(-pident) %>%
        condenseTaxa() %>% # agreement in Phylogeny
      paste(., collapse = "%")

       

    }else{
      # If there are no matches, better than the cutoff, then keep everything
      
    .x %>%
        select(-pident) %>%
    condenseTaxa() %>%
       paste(., collapse = "%")
      }
  }
  }
  
  # Collapse all the taxa data separatated by %, como queda feo para leer en excel lo modificamos con # PERO es un lio dejarlo sin el % porq deja la table separada sin heads asi que mejor dejarlo como esta y luego en R separar las columnas por % y asignarles nombres
  
  )) %>%
  select(qseqid, consensus) %>%
  unnest(consensus)}

```


## Apply the function to our data

```{r}
all.lineages.long <- read_csv("../data/taxonomies_matches.csv")
all.lineages.long %>%
  rename(name = 2) %>%
  group_by(taxID, rank) %>%
  tally() %>%
  filter (n>1) %>%
  distinct(taxID) %>%
  inner_join(BLAST_results %>% mutate (taxID = as.numeric(staxid))) %>%
  ungroup() %>%
  distinct(qseqid) %>%
  inner_join(BLAST_results)

## From all this, we can dismiss the two staxid for which we have two filogenies
  all.lineages.long %>%
  # rename(name = 2) %>%
  group_by(taxID, rank) %>%
  tally() %>%
  filter (n==1) %>%
    select(-n) %>%
    inner_join(all.lineages.long,.) %>%

pivot_wider(names_from = rank, values_from = 2)-> readytojoin
```
3731 taxID and 8 collumns
lo que no tengo lineage information se remueve con el inner_join. Perdemos la informacion de algunos
BLAST_results %>% anti_join(readytojoin,by=c("staxid"="taxID")) -> para tener la info de los sin lineage information
Luego se puede explorar la informacion que hay dentro de los sin lineage BLAST_results %>% anti_join(readytojoin,by=c("staxid"="taxID")) %>% distinct(staxid) y me dio 141 hits. Que son muestras con problemas en el taxaid que tienen ; y contamos los casos y la mayoria se encuentran con dos o tres opciones de id lineage info por eso nos da error en el consensus ya que tiene diferencias.


```{r}

BLAST_results %>%
  mutate(taxID = as.numeric(staxid)) %>%
  inner_join(readytojoin %>%
  mutate(taxID = as.numeric(taxID))) -> input

# Store the max pident observed per qseqid

input |> 
  group_by(qseqid) |> 
  summarise (max_pident = max(pident)) -> max_pidents


custom.lca(input,95) -> finalids

finalids %>%
  separate(consensus, into = c("kingdom", "phylum", "class", "order", "family", "genus", "species"), sep = "%") %>%
   inner_join(max_pidents) |> 
  rename(Hash = 1) %>%
 
  write_csv("../pipeline_output/hash_classified_95_100.csv")
```

```{r}
thresholds <- list(100, 97, 95)
map(thresholds, ~custom.lca( input, .x)) -> ids_thresholds

output.summary <- function(tibble){
  
  tibble %>% 
    rename(Hash =1) %>% 
    separate(consensus, into = c("kingdom", "phylum", "class", "order", "family", "genus", "species"), sep = "%") %>% 
    # pivot_longer(-Hash, names_to = "rank", values_to = "name") %>% 
    mutate (final_rank = case_when(species != "NA" ~ "species",
                                   genus != "NA" ~ "genus",
                                   TRUE          ~ "Worse")) %>% 
    group_by(final_rank) %>% 
    tally()
}

map(ids_thresholds, output.summary)

## Write_ 100 to spp
map(list("95"=finalids), output.summary)

ids_thresholds[[1]] %>% 
  rename(Hash =1) 
```

